---
title: "AdmissionEDA"
author: "Maha Mapara"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(caret)
library(lubridate)
library(lattice)
library(gridExtra)
library(GGally)
library(ISLR)
library(tidyr)

install.packages("naniar")
library(naniar)
```


```{r}
library(readxl)
MassMutualTestJumpStartDataAnalysis20191010 <- read_excel("MassMutualTestJumpStartDataAnalysis20191010.xlsx")
```

Response: Application
Explanatory Variables:
Columns of interest: 10-18, 21(race international), 29(first generation), 22-27(race), 39(legacy), 43(sib applied), 44(SAT considered), 48(ACT considered)
55 (CompStandizedTestPercentile), 62-66(% of interests in that academic area(should add later)), 67(recruited athlete), 74(HS US Zipcode)
85-85 (measures of engagement), 86(total engagement), 97(total_income), 98(sfs family contribution), 103(financial_aid, 116(21centmerit), 117(posse_applicant))

132-135: 132-PreApplicationMessagesOpen; 133-PreApplicationMessageSent; 134-PostApplicationMessageSent; 135-PostApplicationsMessagesOpen


```{r}
admissions <- MassMutualTestJumpStartDataAnalysis20191010
```


```{r}
#rename columns
names(admissions)[14] <- "accept_ED_as_RD"
names(admissions)[18] <- "final_app_rating"
names(admissions)[21] <- "race_international"
names(admissions)[22] <- "hispanic"
names(admissions)[23] <- "american_indian"
names(admissions)[24] <- "asian_american"
names(admissions)[25] <- "african_american"
names(admissions)[26] <- "pacific_islander_american"
names(admissions)[27] <- "white_american"
names(admissions)[29] <- "first_gen"
names(admissions)[39] <- "legacy"
names(admissions)[43] <- "sib_applied"
names(admissions)[44] <- "sat_considered"
names(admissions)[48] <- "act_considered"
names(admissions)[56] <- "toefl"
names(admissions)[57] <- "ielts"
names(admissions)[61] <- "acad_int_count"
names(admissions)[67] <- "recruited_athlete"
names(admissions)[74] <- "hs_us_zipcode"
names(admissions)[80] <- "hs_gpa"
names(admissions)[96] <- "total_engagement"
names(admissions)[97] <- "total_income"
names(admissions)[98] <- "sfs_fam_cont"
names(admissions)[103] <- "financial_aid"
names(admissions)[104] <- "orig_need"
names(admissions)[116] <- "cent_merit"
names(admissions)[117] <- "posse_applicant"
names(admissions)[118] <- "posse_admit"
names(admissions)[136] <- "deposit_indicator"





```

```{r}
str(admissions)
```

```{r}
b <- admissions$Admit

count0 <- length(which(b == 1))

count0

b <- admissions$Application

count0 <- length(which(b == 1))

count0
```

Boxplots
```{r}
#boxplot(admissions$Admit)
# boxplot(admissions$total_income)
# boxplot(admissions$sfs_fam_cont)
# boxplot(admissions$orig_need)
# boxplot(admissions$hs_gpa)

# b1 <- boxplot(admissions$Application,
#               main = "Application",
#               xlab = "Applied or not",
#               col = "orange",
#               border = "brown",
#               horizontal = TRUE,
#               notch = TRUE
#               )

# b2 <- boxplot(admissions$Admit,
#               main = "Admitted/Accepted",
#               xlab = "Admitted or not",
#               col = "orange",
#               border = "brown",
#               horizontal = TRUE,
#               notch = TRUE
#               )

b3 <- boxplot(admissions$total_income,
              main = "Total Family Income",
              xlab = "Family Income",
              col = "orange",
              border = "brown",
              horizontal = TRUE,
              notch = TRUE
              )

b4 <- boxplot(admissions$sfs_fam_cont,
              main = "SFS Family Contribution",
              xlab = "Family Contribution",
              col = "orange",
              border = "brown",
              horizontal = TRUE,
              notch = TRUE
              )

b5 <- boxplot(admissions$hs_gpa,
              main = "Final High School GPA",
              xlab = "High School GPA",
              col = "orange",
              border = "brown",
              horizontal = TRUE,
              notch = TRUE
              )

b6 <- boxplot(admissions$orig_need,
              main = "Original Need spread for applicants in 2017",
              xlab = "Original Need",
              col = "orange",
              border = "brown",
              horizontal = TRUE,
              notch = TRUE
              )

b3
b4
b5
b6
```

```{r}
h1 <- hist(admissions$hs_gpa, col = "green", breaks = 100)

h2 <- hist(admissions$orig_need, col = "green", breaks = 100)

h3 <- hist(admissions$total_income, col = "green", breaks = 100)

h4 <- hist(admissions$sfs_fam_cont, col = "green", breaks = 100)

h1
h2
h3
h4
```


```{r}
#install.packages(Amelia)
library(Amelia)
library(mlbench)

missmap(admissions, col = c("blue", "red"), legend = FALSE)
```


```{r}
# pairs(~total_income+sfs_fam_cont+hs_gpa,aes(color=Admit), data=admissions)
# 
# vars_to_use <- c("total_income", "sfs_fam_cont", "hs_gpa", "orig_need", data = admissions)
# ggpairs(vars_to_use, aes(color = Admit))

```

```{r}
#plot(wt, mpg, main="Scatterplot Example"
     
# pairs(~Application+race_international+first_gen+legacy,data=admissions)
```

```{r}
#use following columns for pair plots:
#21, 29, 10-18, 39, 43, 44, 48, 55, 67, 86, 97, 98, 103, 116, 117

# vars_to_use1 <- c("Application", "accept_ED_as_RD", "race_international", "first_gen", "legacy", "sib_applied", "sat_considered", "act_considered", "recruited_athlete")
# # 
# ggpairs(vars_to_use1, data = admissions)
# 
# 
# 
# vars_to_use2 <- c("application", "inquiry","total_engagement", "total_income", "sfs_fam_cont", "financial_aid", "cent_merit", "posse_applicant", "CompStandizedTestPercentile")
# 
# ggpairs(vars_to_use2, data = admissions)
# 
# 
# #"hs_us_zipcode" Column 'hs_us_zipcode' has more levels (14669) than the threshold (15) allowed. Please #remove the column or increase the 'cardinality_threshold' parameter. Increasing the #cardinality_threshold may produce long processing times
```

```{r}
pairs(vars_to_use, col = admissions$Admit)
```

####remove all NA rows in race background: remove cell where 0.28 and 0.05 hispanic, 0.68 with white, etc 

#take off acad_int_count for now cuz in some cases col 2 is only pop and counted as two interest counts-- where is data coming from

```{r}

#library(naniar)

####remove all NA rows in race background: remove cell where 0.28 and 0.05 hispanic, 0.68 with white, american indian with 0.04, asian with 0.17, african with 0.11, pacific islander with 0.01.
admissions <- admissions %>%
  replace_with_na(replace = list(hispanic= c(0.28, 0.05),
                             white_american = c(0.68),
                             american_indian = c(0.04),
                             asian_american = c(0.17),
                             african_american = c(0.11),
                             pacific_islander_american = c(0.01)))
View(admissions)

##are we not accounting for students who have more than once race?
#they go into color even if white and another race(s)
```
#make all dom POC in one column:hispanic, american_indian, asian_american, african_american, pacific_islander_american

```{r}
admissions$dom_stu_of_colour <- admissions$hispanic+ admissions$american_indian+ admissions$asian_american+ admissions$african_american+ admissions$pacific_islander_american

View(admissions)
```
#race colunms: dom_stu_of_colour, white_american, race_international
#rename white_american, race_international:
```{r}
names(admissions)[21] <- "int_stu"
names(admissions)[27] <- "dom_white"
```


after this go to interpretation of odds ratios etc and include dummy var dom_color in models too.


references:

https://cran.r-project.org/web/packages/naniar/readme/README.html

https://cran.r-project.org/web/packages/naniar/vignettes/replace-with-na.html

https://cran.r-project.org/web/packages/fastDummies/vignettes/making-dummy-variables.html
