---
title: "Cleaning up"
author: "Maha Mapara"
date: "11/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(caret)
library(lubridate)
library(lattice)
library(gridExtra)
library(GGally)
library(ISLR)
library(tidyr)
library(readxl)

#install.packages("naniar")
library(naniar)
```


```{r}
MassMutualTestJumpStartDataAnalysis20191010 <- read_excel("MassMutualTestJumpStartDataAnalysis20191010.xlsx")
```

```{r}
admissions <- MassMutualTestJumpStartDataAnalysis20191010
```


```{r}
#rename columns
names(admissions)[11] <- "application"

names(admissions)[14] <- "accept_ED_as_RD"
names(admissions)[18] <- "final_app_rating"

names(admissions)[19] <- "citizenship"

names(admissions)[21] <- "int_stu"
names(admissions)[22] <- "hispanic"
names(admissions)[23] <- "american_indian"
names(admissions)[24] <- "asian_american"
names(admissions)[25] <- "african_american"
names(admissions)[26] <- "pacific_islander_american"
names(admissions)[27] <- "dom_white"

names(admissions)[29] <- "first_gen"

names(admissions)[39] <- "legacy"
names(admissions)[43] <- "sib_applied"

names(admissions)[44] <- "sat_considered"
names(admissions)[45] <- "sat_comp"
names(admissions)[46] <- "sat_verbal"
names(admissions)[47] <- "sat_math"

names(admissions)[48] <- "act_considered"
names(admissions)[49] <- "act_comp"
names(admissions)[50] <- "act_math"
names(admissions)[51] <- "act_reading"
names(admissions)[52] <- "act_english"
names(admissions)[53] <- "act_writing"
names(admissions)[54] <- "act_science"

names(admissions)[55] <- "comp_standardized_test_percentile"

names(admissions)[56] <- "toefl"
names(admissions)[57] <- "ielts"

names(admissions)[61] <- "acad_int_count"

names(admissions)[62] <- "fine_arts"
names(admissions)[63] <- "humanities"
names(admissions)[64] <- "languages"
names(admissions)[65] <- "stem"
names(admissions)[66] <- "social_sciences"

names(admissions)[67] <- "recruited_athlete"

names(admissions)[70] <- "hs_name"
names(admissions)[72] <- "hs_geomarket"
names(admissions)[73] <- "hs_state"
names(admissions)[74] <- "hs_us_zipcode"
names(admissions)[75] <- "hs_country"
names(admissions)[76] <- "hs_grad_year"
  

names(admissions)[80] <- "hs_gpa"

names(admissions)[81] <- "home_geomarket"
names(admissions)[82] <- "home_city"
names(admissions)[83] <- "home_state"
names(admissions)[84] <- "home_country"

names(admissions)[85] <- "campus_tour"
names(admissions)[86] <- "info_session"
names(admissions)[87] <- "college_fair"
names(admissions)[88] <- "sisters_event"
names(admissions)[89] <- "attended_focus_session"
names(admissions)[90] <- "school_visit_day"
names(admissions)[91] <- "third_party_interview"
names(admissions)[92] <- "community_org_visit"
names(admissions)[93] <- "campus_visit_day"
names(admissions)[94] <- "gathering_event"
names(admissions)[95] <- "student_lunch_event"
names(admissions)[96] <- "total_engagement"

names(admissions)[97] <- "total_income"
names(admissions)[98] <- "sfs_fam_cont"
names(admissions)[103] <- "financial_aid"
names(admissions)[104] <- "orig_need"
names(admissions)[116] <- "cent_merit"
names(admissions)[117] <- "posse_applicant"
names(admissions)[118] <- "posse_admit"

names(admissions)[132] <- "pre_app_mess_open"
names(admissions)[133] <- "pre_app_mess_sent"
names(admissions)[134] <- "PostApplicationMessagesSent"

names(admissions)[136] <- "deposit_indicator"
```


```{r}

#library(naniar)

####remove all NA rows in race background: remove cell where 0.28 and 0.05 hispanic, 0.68 with white, american indian with 0.04, asian with 0.17, african with 0.11, pacific islander with 0.01.
admissions <- admissions %>%
  replace_with_na(replace = list(hispanic= c(0.28, 0.05),
                             dom_white = c(0.68),
                             american_indian = c(0.04),
                             asian_american = c(0.17),
                             african_american = c(0.11),
                             pacific_islander_american = c(0.01)))

##are we not accounting for students who have more than once race?
#they go into color even if white and another race(s)
```

#make all dom POC in one column:hispanic, american_indian, asian_american, african_american, pacific_islander_american

```{r}
admissions$dom_stu_of_colour <- admissions$hispanic+ admissions$american_indian+ admissions$asian_american+ admissions$african_american+ admissions$pacific_islander_american

```

subset admissions data frame to a smaller one with all variable pertaining to application (as outlined in google doc)

```{r}
application <- admissions[c(11,19,137,27,21:26,45:47,49:57,62:67,70,72:76,81:96,133,132)]
View(application)
```

Descriptive Stats-all
```{r}
summary(application)
```


```{r}
test_stats <- application %>% 
  filter(application == 1)
nrow(test_stats)

test_stats <- application %>% 
  filter(application == 1 & recruited_athlete == 1)
nrow(test_stats)

test_stats <- application %>% 
  filter(application == 1 & recruited_athlete == 0)
nrow(test_stats)
#recruited_athlete <- application %>%
#   select(recruited_athlete)
```


```{r}
(1271/22017)*100

(20746/22017)*100
```

So out of the 22017 students that applie to Mount Holyoke, 5.77% were recruited athletes, while 94.23% were not recruited athletes.

```{r}
test_stats <- application %>% 
  filter(application == 0)
nrow(test_stats)

test_stats <- application %>% 
  filter(application == 0 & recruited_athlete == 1)
nrow(test_stats)

test_stats <- application %>% 
  filter(application == 0 & recruited_athlete == 0)
nrow(test_stats)

```

Out of the 340520 stduents that didn't apply, none were recruited athletes.

```{r}
test_stats <- application %>% 
  filter(application == 1)
nrow(test_stats)

summary(test_stats)
```

SAT and ACT score spread (means are above):

```{r}
qplot(application$sat_comp, geom="histogram")

qplot(application$act_comp, geom="histogram")
```

Total engagement:

-using logistic regression to see which ones are significant

```{r}
campus_tour <- glm(application ~ campus_tour, data = application, family = binomial)
summary(campus_tour)
```

```{r}
info_session <- glm(application ~ info_session, data = application, family = binomial)
summary(info_session)
```

```{r}
college_fair <- glm(application ~ college_fair, data = application, family = binomial)
summary(college_fair)
```

```{r}
sisters_event <- glm(application ~ sisters_event, data = application, family = binomial)
summary(sisters_event)
```

```{r}
attended_focus_session <- glm(application ~ attended_focus_session, data = application, family = binomial)
summary(attended_focus_session)
```

```{r}
school_visit_day <- glm(application ~ school_visit_day, data = application, family = binomial)
summary(school_visit_day)
```

```{r}
third_party_interview <- glm(application ~ third_party_interview, data = application, family = binomial)
summary(third_party_interview)
```

```{r}
comm_org_visit <- glm(application ~ community_org_visit, data = application, family = binomial)
summary(comm_org_visit)
```

```{r}
campus_visit_day <- glm(application ~ campus_visit_day, data = application, family = binomial)
summary(campus_visit_day)
```

```{r}
gathering_event <- glm(application ~ gathering_event, data = application, family = binomial)
summary(gathering_event)
```

```{r}
student_lunch_event <- glm(application ~ student_lunch_event, data = application, family = binomial)
summary(student_lunch_event)
```

```{r}
qplot(test_stats$total_engagement, geom="histogram", bins = 30)

ggplot(test_stats, aes(x=total_engagement, y=application)) + 
  geom_point()

ggplot(test_stats, aes(x=application, y=total_engagement)) + 
  geom_boxplot()
```

High School
```{r}
test_stats <- application %>% 
  filter(application == 1)
hs <- application[c(1,29,30)]
hs <- hs %>%
  filter(application == 1)
View(hs)
```

#Transform all to lowercase, change '&' to 'and' , 'sch' to 'school'for hs_name
```{r}
test_stats$hs_name <- tolower(test_stats$hs_name)

chartr(old = "&", new = "and", test_stats$hs_name)
chartr(old = "sch", new = "school", test_stats$hs_name)
chartr(old = "karchi", new = "karachi", test_stats$hs_name)
```

#tells me how many times a particlar HS appears
```{r}
count_for_each_hs <- table(test_stats$hs_name)
#View(count_for_each_hs)
```


```{r}
summary(count_for_each_hs)
max(count_for_each_hs)
min(count_for_each_hs)

```

```{r}
count_for_each_hs <- as.data.frame(count_for_each_hs) #convert to a data frame

#change col names
names(count_for_each_hs)[1] <- "hs_name"
names(count_for_each_hs)[2] <- "num_of_applicants"

```



```{r}
#schools with less than 15 applicants
hs_less_15 <- count_for_each_hs %>% 
  filter(num_of_applicants < 15)


#schools with 15-20 applicants
hs_15_20 <- count_for_each_hs %>% 
  filter(num_of_applicants >= 15 & num_of_applicants<= 20)

#schools with greater than 20 and less than 30
hs_20_30 <- count_for_each_hs %>% 
  filter(num_of_applicants > 20 & num_of_applicants <= 30)

#school with greater than 30 applicants
hs_over_30 <- count_for_each_hs %>% 
  filter(num_of_applicants > 30)

#total number/count and summary in each of the three frequency categories
nrow(hs_less_15)
summary(hs_less_15)

nrow(hs_15_20)
summary(hs_15_20)

nrow(hs_20_30)
summary(hs_20_30)

nrow(hs_over_30)
summary(hs_over_30)
```

do a bar graph with <15, 15-20, 21-30, >30

for this: data frame will have 3 columns= hs_name, no. of apps, app. threshold (change the counts_for_each_hs df)

```{r}
# count_for_each_hs$num_app_thresholds <- NA
# 
# #if num_of_applicants < 15, num_app_thresholds should be less than 15
# for (i in (1 : 4982)){
#   
#   if (count_for_each_hs$num_of_applicants < 15){
#     count_for_each_hs$num_app_thresholds <- 'less than 15'
#     }else if(count_for_each_hs$num_of_applicants >= 15 & count_for_each_hs$num_of_applicants <= 20){
#       count_for_each_hs$num_app_thresholds <- '15 to 20 applicants'
#       }else if(count_for_each_hs$num_of_applicants > 20 & count_for_each_hs$num_of_applicants <= 30){
#         count_for_each_hs$num_app_thresholds <- '21 to 30 applicants'
#         }else{
#           count_for_each_hs$num_app_thresholds <- 'greater than 30 applicants'
#       }
#     
# }
  
#error is that the condition has length > 1 and only the first element will be used. so it's putting less than 15 for all, even when not < 15

#figure out how to fix this

#in the mean time: for each app threshold df, add the new column of NAs, and assign those strings. then r bind all the df's

hs_less_15$num_app_thresholds <- NA
hs_less_15$num_app_thresholds <- 'less than 15'

hs_15_20$num_app_thresholds <- NA
hs_15_20$num_app_thresholds <- '15 to 20 applicants'

hs_20_30$num_app_thresholds <- NA
hs_20_30$num_app_thresholds <- '21 to 30 applicants'
  
hs_over_30$num_app_thresholds <- NA
hs_over_30$num_app_thresholds <- 'greater than 30 applicants'

count_with_thresholds <- rbind(hs_less_15, hs_15_20)
count_with_thresholds <- rbind(count_with_thresholds, hs_20_30)
count_with_thresholds <- rbind(count_with_thresholds, hs_over_30)

View(count_with_thresholds)

```




```{r}
#ggplot(data = count_for_each_hs, aes(x = num_of_applicants, y = hs_name)) + geom_col()

tail(count_for_each_hs)

```

```{r}
ggplot(count_with_thresholds, aes(num_app_thresholds,num_of_applicants))+
  geom_col()
```

Going back to SAT scores
```{r}
test_stats <- application %>% 
  filter(application == 1)#to make sure app == 1

SAT_count <- table(test_stats$sat_comp)
SAT_count <- as.data.frame(SAT_count)

names(SAT_count)[1] <- "sat_score"
names(SAT_count)[2] <- "num_of_applicants"

#sat_score was a factor and changed it to number
SAT_count$sat_score <- as.numeric(levels(SAT_count$sat_score))[SAT_count$sat_score]

lowSAT <- SAT_count %>%
  filter(sat_score < 1400)

highSAT <- SAT_count %>%
  filter(sat_score >= 1400)

lowSAT$label <- NA
lowSAT$label <- '< 1400'

highSAT$label <- NA
highSAT$label <- '>= 1400'

SAT_count <- rbind(lowSAT, highSAT)

ggplot(SAT_count, aes(label,num_of_applicants))+
  geom_col()

#View(SAT_count)
# SAT <- test_stats %>%
#   filter(sat_comp >= 1400)

```

```{r}
pre_app_messages <- test_stats %>%
  select(application, pre_app_mess_sent, pre_app_mess_open)

mess_sent_count <- table(test_stats$pre_app_mess_sent)
mess_sent_count <- as.data.frame(mess_sent_count)
names(mess_sent_count)[1] <- "num_of_messages_sent"
names(mess_sent_count)[2] <- "frequency"

mess_open_count <- table(test_stats$pre_app_mess_open)
mess_open_count <- as.data.frame(mess_open_count)
names(mess_open_count)[1] <- "num_of_messages_opened"
names(mess_open_count)[2] <- "frequency"
##
mess_sent_count$num_of_messages_sent <-
  as.numeric(levels(mess_sent_count$num_of_messages_sent))[mess_sent_count$num_of_messages_sent]

mess_open_count$num_of_messages_opened <-
  as.numeric(levels(mess_open_count$num_of_messages_opened))[mess_open_count$num_of_messages_opened]

##

mess_sent_lesseq7 <- mess_sent_count %>%
  filter(num_of_messages_sent <= 7)

mess_sent_8to15 <- mess_sent_count %>%
  filter(num_of_messages_sent > 7 & num_of_messages_sent <= 15)

mess_sent_16to23 <- mess_sent_count %>%
  filter(num_of_messages_sent > 15 & num_of_messages_sent <= 23)

mess_sent_gretereq24 <-  mess_sent_count %>%
  filter(num_of_messages_sent >= 24)

mess_sent_lesseq7$label <- NA
mess_sent_lesseq7$label <- 'less than or equal to 7'

mess_sent_8to15$label <- NA
mess_sent_8to15$label <- '8 to 15'

mess_sent_16to23$label <- NA
mess_sent_16to23$label <- '16 to 23'

mess_sent_gretereq24$label <- NA
mess_sent_gretereq24$label <- 'greater than or equal to 24'

mess_sent_count <- rbind(mess_sent_lesseq7, mess_sent_8to15)
mess_sent_count <- rbind(mess_sent_count, mess_sent_16to23)
mess_sent_count <- rbind(mess_sent_count, mess_sent_gretereq24)
##

mess_open_less7 <- mess_open_count %>%
  filter(num_of_messages_opened < 7)

# mess_open_2to7 <- mess_open_count %>%
#   filter(num_of_messages_opened > 1 & num_of_messages_opened <= 7)

mess_open_7to15 <- mess_open_count %>%
  filter(num_of_messages_opened >= 7 & num_of_messages_opened <= 15)

mess_open_morethan15 <- mess_open_count %>%
  filter(num_of_messages_opened > 15)

mess_open_less7$label <- NA
mess_open_less7$label <- 'less than 7'

mess_open_7to15$label <- NA
mess_open_7to15$label <- '7 to 15'

mess_open_morethan15$label <- NA
mess_open_morethan15$label <- 'greater than 15'

mess_open_count <- rbind(mess_open_less7, mess_open_7to15)
mess_open_count <- rbind(mess_open_count, mess_open_morethan15)

#mess_open_count <- rbind(mess_open_count, mess_open_morethan15)
##

ggplot(mess_sent_count, aes(label,num_of_messages_sent))+
  geom_col()

ggplot(mess_open_count, aes(label,num_of_messages_opened))+
  geom_col()

```

for ineractions and non-linearity
```{r}
#use for interactions

# ggplot(data = bike, mapping = aes(x = mnth, y = cnt, color = factor(workingday))) + 
#   geom_point() +
#   geom_smooth()
# 
# ggplot(data = bike, mapping = aes(x = factor(mnth), y = cnt, color = factor(workingday))) + 
#   geom_boxplot()

```


```{r}

#install.packages("epiDisplay")

# library(epiDisplay)
# 
# par(mar=c(1,1,1,1))
# 
# tab1(test_stats$hs_name, sort.group = "decreasing", cum.percent = TRUE)
```

```{r}
# install.packages("summarytools")
# library(summarytools)
# summarytools::freq(test_stats$hs_name, order = "freq")
```


##
References:
https://www.statmethods.net/management/subset.html

https://www.statmethods.net/stats/descriptives.html

https://dabblingwithdata.wordpress.com/2017/12/20/my-favourite-r-package-for-frequency-tables/

https://bradleyboehmke.github.io/tutorials/manipulation_base_r

http://rprogramming.net/r-data-manipulation/

