---
title: "Enrollment clean up and plots"
author: "Maha Mapara"
date: "2/14/2020"
output: html_document
---

Summary and commit messgae: 

Renamed columns, created dom_stu_of_color(not working), created enrollment dataframe with all variables imp for enrollment (only RD and up to 2019), created enrolled with Enroll == 1, many plots for correlations


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

logistic regression, high yield, low yield, insitutional grant/1000

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(caret)
library(lubridate)
library(lattice)
library(gridExtra)
library(GGally)
library(ISLR)
library(tidyr)
library(readxl)

#install.packages("naniar")
library(naniar)
```

```{r}
#MassMutualTestJumpStartDataAnalysis20191010 <- read_excel("MassMutualTestJumpStartDataAnalysis20191010.xlsx")
```

```{r}
#admissions <- MassMutualTestJumpStartDataAnalysis20191010

admissions <-InqMMJumpStartDataAnalysis20200128_1_
```


```{r}
#rename columns
names(admissions)[9] <- "original_app_plan"
names(admissions)[11] <- "application"

names(admissions)[14] <- "accept_ED_as_RD"
names(admissions)[18] <- "final_app_rating"

names(admissions)[19] <- "citizenship"

names(admissions)[21] <- "int_stu"
names(admissions)[22] <- "hispanic"
names(admissions)[23] <- "american_indian"
names(admissions)[24] <- "asian_american"
names(admissions)[25] <- "african_american"
names(admissions)[26] <- "pacific_islander_american"
names(admissions)[27] <- "dom_white"

names(admissions)[29] <- "first_gen"

names(admissions)[39] <- "legacy"
names(admissions)[43] <- "sib_applied"

names(admissions)[44] <- "sat_considered"
names(admissions)[45] <- "sat_comp"
names(admissions)[46] <- "sat_verbal"
names(admissions)[47] <- "sat_math"

names(admissions)[48] <- "act_considered"
names(admissions)[49] <- "act_comp"
names(admissions)[50] <- "act_math"
names(admissions)[51] <- "act_reading"
names(admissions)[52] <- "act_english"
names(admissions)[53] <- "act_writing"
names(admissions)[54] <- "act_science"

names(admissions)[55] <- "comp_standardized_test_percentile"

names(admissions)[56] <- "toefl"
names(admissions)[57] <- "ielts"

names(admissions)[61] <- "acad_int_count"

names(admissions)[62] <- "fine_arts"
names(admissions)[63] <- "humanities"
names(admissions)[64] <- "languages"
names(admissions)[65] <- "stem"
names(admissions)[66] <- "social_sciences"

names(admissions)[67] <- "recruited_athlete"

names(admissions)[70] <- "hs_name"
names(admissions)[72] <- "hs_geomarket"
names(admissions)[73] <- "hs_state"
names(admissions)[74] <- "hs_us_zipcode"
names(admissions)[75] <- "hs_country"
names(admissions)[76] <- "hs_grad_year"
  

names(admissions)[80] <- "hs_gpa"

names(admissions)[81] <- "home_geomarket"
names(admissions)[82] <- "home_city"
names(admissions)[83] <- "home_state"
names(admissions)[84] <- "home_country"

names(admissions)[85] <- "campus_tour"
names(admissions)[86] <- "info_session"
names(admissions)[87] <- "college_fair"
names(admissions)[88] <- "sisters_event"
names(admissions)[89] <- "attended_focus_session"
names(admissions)[90] <- "school_visit_day"
names(admissions)[91] <- "third_party_interview"
names(admissions)[92] <- "community_org_visit"
names(admissions)[93] <- "campus_visit_day"
names(admissions)[94] <- "gathering_event"
names(admissions)[95] <- "student_lunch_event"
names(admissions)[96] <- "total_engagement_pre_app"

names(admissions)[97] <- "total_income"
names(admissions)[98] <- "sfs_fam_contribution"
names(admissions)[102] <- "sfs_inst_grant"
names(admissions)[103] <- "financial_aid"
names(admissions)[104] <- "orig_need"
names(admissions)[108] <- "total_inst_grant_cofhe"
names(admissions)[112] <- "pell_grant_offered"
names(admissions)[114] <- "loan_flag_cofhe"
names(admissions)[115] <- "loan_value_cofhe"
names(admissions)[116] <- "21_cent_merit"
names(admissions)[117] <- "posse_applicant"
names(admissions)[118] <- "posse_admit"
names(admissions)[119] <- "total_inst_grant" #need and non-need
names(admissions)[120] <- "total_grants_awarded" #all grants from all sources
names(admissions)[121] <- "total_loans_awarded"

names(admissions)[124] <- "attending_instead"
names(admissions)[125] <- "native_language"

names(admissions)[128] <- "deposit_indicator"
names(admissions)[131] <- "all_mess_sent"
names(admissions)[132] <- "all_mess_open"
names(admissions)[133] <- "pre_app_mess_open"
names(admissions)[134] <- "pre_app_mess_sent"
names(admissions)[135] <- "post_app_mess_sent"
names(admissions)[136] <- "post_app_mess_open"

names(admissions)[138] <- "athletic_interest"
names(admissions)[139] <- "application_created"
names(admissions)[140] <- "deposit_received_date"
names(admissions)[143] <- "institute_attending_instead"
names(admissions)[144] <- "2_or_4_year"
names(admissions)[145] <- "public_or_private"
names(admissions)[149] <- "womens_college"
names(admissions)[150] <- "lib_arts_college"
names(admissions)[157] <- "girls_only_hs"

names(admissions)[158] <- "mhc_preview_day"
names(admissions)[159] <- "mhc_showcase_event"
```


```{r}

#library(naniar)

####remove all NA rows in race background: remove cell where 0.28 and 0.05 hispanic, 0.68 with white, american indian with 0.04, asian with 0.17, african with 0.11, pacific islander with 0.01.
admissions <- admissions %>%
  replace_with_na(replace = list(hispanic= c(0.28, 0.05),
                             dom_white = c(0.68),
                             american_indian = c(0.04),
                             asian_american = c(0.17),
                             african_american = c(0.11),
                             pacific_islander_american = c(0.01)))

##are we not accounting for students who have more than once race?
#they go into color even if white and another race(s)
```

#make all dom POC in one column:hispanic, american_indian, asian_american, african_american, pacific_islander_american

```{r}
admissions$dom_stu_of_colour <- admissions$hispanic+ admissions$american_indian+ admissions$asian_american+ admissions$african_american+ admissions$pacific_islander_american

```

```{r}
race <- c("int_stu", "dom_white","hispanic","american_indian","asian_american","african_american","pacific_islander_american" )

racial_categories <- admissions[race]
```

#add financial aid stuff like inst grants etc.
```{r}
for_new <- c("FallYear","original_app_plan", "application", "Admit","Enroll", "accept_ED_as_RD", "final_app_rating",
             
            "native_language", "int_stu", "dom_stu_of_colour", "dom_white",
             
             "first_gen", "legacy", "sib_applied", "recruited_athlete", "athletic_interest","posse_applicant","posse_admit", "girls_only_hs",
            "application_created",
             
             "sat_considered", "act_considered",  "comp_standardized_test_percentile" , "hs_gpa",
            
            "sat_comp","sat_verbal","sat_math","act_comp","act_math","act_reading","act_english","act_writing","act_science","toefl","ielts",
             
             "hs_name", "hs_us_zipcode","HighYieldFeederHS" ,"LowYieldFeederHS", "HighYieldZIPTrack", "LowYieldZIPTrack", 
            #highyieldfeederhs was >15 for 3 years or >15 admits
            #lowyieldfeederhs was less than 15 or more than 15 but no admits in the 3 years
            #ziptrack: first 3 characters of hs zipcode
             
             "financial_aid", "orig_need","sfs_fam_contribution","sfs_inst_grant", "total_inst_grant_cofhe", 
             "pell_grant_offered","loan_flag_cofhe","loan_value_cofhe","21_cent_merit" ,"total_inst_grant", "total_grants_awarded", "total_loans_awarded", "SelfhelpWorkFlagCOFHE",
             
             "acad_int_count",
            
            "campus_tour","info_session","college_fair","sisters_event","attended_focus_session","school_visit_day","third_party_interview","community_org_visit","campus_visit_day","gathering_event","student_lunch_event","total_engagement_pre_app" ,"mhc_preview_day", "mhc_showcase_event", 
            
            "all_mess_sent", "all_mess_open", "pre_app_mess_open", "pre_app_mess_sent", "post_app_mess_open","post_app_mess_sent",
             
             
            "deposit_indicator","deposit_received_date",
            
            "attending_instead","2_or_4_year","public_or_private","womens_college","lib_arts_college")

enrollment <- admissions[for_new] #68 variables. original had 166

```

#exclude ED1 and ED2 from plots and applications in 2020
```{r}
enrollment <- enrollment %>%
  filter(FallYear != 2020)

enrollment <- enrollment %>%
  filter(original_app_plan == "RD")
```


```{r}
#total grants awarded

ggplot(data = enrollment, mapping = aes(x = total_grants_awarded, y = Enroll)) + 
  geom_point() +
  geom_smooth()

#total loans awarded

ggplot(data = enrollment, mapping = aes(x = total_loans_awarded, y = Enroll)) + 
  geom_point() +
  geom_smooth()

#total_inst_grants

ggplot(data = enrollment, mapping = aes(x = total_inst_grant, y = Enroll)) + 
  geom_point() +
  geom_smooth()

#correlation b/w total grnats and total loans
ggplot(data = enrollment, mapping = aes(x = total_grants_awarded, y = total_loans_awarded, color = factor(Enroll))) + 
  geom_point() 

#correlation b/w sat_verbal and toefl
ggplot(data = enrollment, mapping = aes(x = sat_verbal, y = toefl, color = factor(Enroll))) + 
   geom_point()+
  geom_smooth()

#post application messages sent for enrollment
ggplot(data = enrollment, mapping = aes(x = post_app_mess_sent, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#post application messages opened for enrollment
ggplot(data = enrollment, mapping = aes(x = post_app_mess_open, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#correlation between post app messages sent and opened
ggplot(data = enrollment, mapping = aes(x = post_app_mess_sent, y = post_app_mess_open, color = factor(Enroll))) + 
   geom_point()
 
#all messages opened and enrollment
ggplot(data = enrollment, mapping = aes(x = all_mess_open, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#all opened messages and post messages opened
ggplot(data = enrollment, mapping = aes(x = all_mess_open, y = post_app_mess_open, color = factor(Enroll))) + 
   geom_point()

#all opened messages and post messages sent
ggplot(data = enrollment, mapping = aes(x = post_app_mess_sent, y = all_mess_open, color = factor(Enroll))) + 
   geom_point()+
  geom_smooth()

#toefl and enrollment
ggplot(data = enrollment, mapping = aes(x = toefl, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#act_english and enrollment
ggplot(data = enrollment, mapping = aes(x = act_english, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#act_math and enrollment
ggplot(data = enrollment, mapping = aes(x = act_math, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#act_science and enrollment
ggplot(data = enrollment, mapping = aes(x = act_science, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#act_reading and enrollment
ggplot(data = enrollment, mapping = aes(x = act_reading, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#act_comp and enrollment
ggplot(data = enrollment, mapping = aes(x = act_comp, y = Enroll)) + 
   geom_point()+
  geom_smooth()

#inst grnats and total grants correlation
ggplot(data = enrollment, mapping = aes(x = total_inst_grant, y = total_grants_awarded)) + 
   geom_point()+
  geom_smooth()

#total engagement pre app and enrollment
ggplot(data = enrollment, mapping = aes(x = total_engagement_pre_app, y = Enroll)) + 
   geom_point()+
  geom_smooth()



```
```{r}
#enrolled has only applicants that enrolled at Mount Holyoke

enrolled <- enrollment %>%
  filter(Enroll == 1)

#trend between total grnats and total loans exist even when all obs are enrolled
ggplot(data = enrolled, mapping = aes(x = total_grants_awarded, y = total_loans_awarded)) + 
  geom_point()

#correlation between sat_verbal and toefl for enrolled students
ggplot(data = enrolled, mapping = aes(x = sat_verbal, y = toefl)) + 
  geom_point()+
  geom_smooth()




```





